\documentclass{article}
\usepackage{amsmath}
\usepackage{titling}
\usepackage{float}
\usepackage{graphicx}

\title{Calculus}
\preauthor{}
\author{}
\postauthor{}
\predate{}
\date{}
\postdate{}

\begin{document}

\maketitle

\section{Introduction}
\input{intro.tex}

\section{Derivatives}
\input{derivatives.tex}

\section{Applications Part One}
\input{applicationsp1.tex}

\section{Interlude: Limits}
\input{limits.tex}

\section{Derivatives Strike Back}
\input{derivatives2.tex}

\section{Integrals}
\input{integrals.tex}

\section{Applications Part Two}
\input{applicationsp2.tex}

\section{Differential Equations}
\input{difeqs.tex}

\section{Sequences and Series}

\section{Resources}
Some common single-variable calculus textbooks out there are Larson and Edwards' (more high school oriented) and Spivak's (more college oriented, but I've heard it's very good - unlike Apostol's, which I have read some of and detest with a passion as it's thorough but incredibly dry). Like any book on this list, if you can't find a cheap copy at a used bookstore or a free one through a library, you definitely shouldn't try pirating it through LibGen (gen.lib.rus.ec). Definitely not. Ahem.
\begin{itemize}
    \item Paul's Online Math Notes (I don't think they're great for learning, but they are good for reference and have problems with solutions along with proofs of everything.)
    \item Project Gutenberg has several free calculus books online, including Silvanus Thompson's {\it Calculus Made Easy}, which is very good.
    \item I don't much like Khan Academy's videos personally, but they have a ton of practice problems, and can give you an idea of how much of the whole subject you've mastered.
    \item While I haven't personally read it (it's on my list), I've heard very good things about Tarasov's Calculus - it's also free as a pdf online.
    \item If you get stuck, need help with a problem, or similar, try Math Stack Exchange (some of the questions are also just interesting reading in their own right).
    \item 3blue1brown on YouTube has an excellent series on  how calculus works, with beautiful graphics.
    \item \textit{Burn Math Class} - wonderful explanations of many different parts of math, including calculus, along with satisfying tirades about the tragedy that is the American educational system. Where the example with area in the introduction is from. Frankly, would highly recommend you read this above quite possibly anything else on this list.
\end{itemize}
\end{document}
