\begin{thebibliography}[9]
\bibitem{blubrown} 3Blue1Brown \textit{Essence of Linear Algebra} \texttt{https://www.youtube.com/watch?v=kjBOesZCoqc&list=PLZHQObOWTQDPD3MizzM2xVFitgF8hE_ab} \\3Blue1Brown's linear algebra videos are very good if you want to learn linear algebra.
\bibitem{khanacademy} Khan Academy \textit{Multivariable Calculus} \texttt{https://www.khanacademy.org/math/calculus-home/multivariable-calculus} \\Khan Academy's multivariable calculus series is very good if you want to start learning multivariable calculus.
\bibitem{passcalculus} passcalculus.com \texit{Limits of Functions} \texttt{https://www.youtube.com/watch?v=trAmOJ8pHNc} \\These are good for extra practice problems and another explanation of how limits work. An eleven video series.
\bibitem{mitocw} MIT OCW \textit{Single Variable Calculus} \texttt{https://ocw.mit.edu/courses/mathematics/18-01sc-single-variable-calculus-fall-2010/} \\MIT OCW's online courses are very good. The Calculus I course includes the lectures and (if you have the textbook) all the necessary materials to get going.
\bibitem{paulmath} Prof. Paul Dawkins \textit{Calculus I, II, III} \texttt{http://tutorial.math.lamar.edu/} \\Paul's Online Notes have notes on algebra, precalculus, calculus I, calculus II, and calculus III, as well as practice problems.
\bibitem{apostoli} Apostol \textit{Calculus, Vol. 1 - Single Variable Calculus and Linear Algebra}
\bibitem{apostolii} Apostol \textit{Calculus, Vol. 2 - Multivariable Calculus}
\bibitem{spivak} Spivak \textit{Calculus, 10th Edition}
\bibitem{cartoon} Larry Gonick \textit{The Cartoon Guide to Calculus}
\end{thebibliography{9}